---
title: "nutnet disease"
author: "Katie Schroeder"
date: "2023-01-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## loading in
```{r loading}
#load in data
ks_disease<-read.csv("Stomata_Data_REU_2019_master.csv")
ps_disease<-read.csv("nutnet_fungus_2019_master.csv")
ps_disease_pk<-read.csv("nutnet_fungus_2019_pk_added_col_added.csv")
cover<-read.csv("NutNet_Species_Cover_May30_thru_August.csv")

library(tidyverse)
```

## Basics of each data frame
ks_disease = disease scoring by Megan and Katie for 2019 REU project.
plots: 13 - 24
treatments: all trts of block 2
dates sampled: "6.18.19"   "6.21.2019" "7.10.2019" "7.31.2019"
plants scored:"AGROPYRON REPENS"    "ANDROPOGON GERARDII" "POA PRATENSIS"

ps_disease = disease scoring by Rachel Paseka and Alex Strauss in summer 2019. Lots of manual labor by Evelyn and Hajira and maybe Sam?
plots: 1-60
treatments: all trts all blocks
dates sampled: "6/5/2019" "7/8/2019" "8/6/2019"
plants scored: "Agropyron repens"    "Andropogon gerardii"

cover = cover data by Megan and Katie for 2019 field season.
plots: 1-36, 39, 43, 44, 45, 48, 50, 55, 56, 58, 59
dates: "5.30.2019" "6.18.2019" "6.4.2019"  "7.10.2019" "7.23.2019" "7.24.2019" "7.8.2019" 
treatments: All for blocks 1-3, C, N1, N5, N10 for blocks 4 and 5

## Summaries
```{r summarize}
ps_disease %>%
  count(date,host,plot)
#1-20 replicates per plot per sampling day (most 1-20). Plants were flagged and tracked through time
ps_disease %>%
  count(date,host,trt)
#40-80 replicates per trt per sampling day

ks_disease %>% 
  filter(!is.na(fungal)) %>%
  count(date,plant,trt)
#5-7 replicates per plot per sampling day
```

## General trends
```{r}
#make a total fungal column in alex and rachel's data
ps_disease$fungal_old <- ps_disease$old_rust + ps_disease$old_bs
ps_disease$fungal_new <- ps_disease$new_rust + ps_disease$new_bs

library(ggpubr)
hist(ps_disease$old_rust,breaks=50)
hist(ps_disease$old_bs,breaks=50)

hist(ps_disease$new_rust,breaks=50)
#old rust is most frequently not 0

ps_disease_pk$trt<-factor(ps_disease$trt,levels=c("C","C_H","N1","N5","N10","P","K","PK","NK","NP","NPK","NPK_H"))
ggerrorplot(filter(ps_disease,trt=="C"|trt=="N1"|trt=="N5"|trt=="N10"|trt=="PK"|trt=="NPK"),x="trt",y="old_rust",desc_stat = "mean_sd",facet.by=c("date","host"))
ggerrorplot(filter(ps_disease,trt=="C"|trt=="N1"|trt=="N5"|trt=="N10"|trt=="PK"|trt=="NPK"),x="trt",y="new_rust",desc_stat = "mean_sd",facet.by=c("date","host"))

#invert damage
ggerrorplot(filter(ps_disease,trt=="C"|trt=="N1"|trt=="N5"|trt=="N10"|trt=="PK"|trt=="NPK"),x="trt",y="old_invert",desc_stat = "mean_sd",facet.by=c("date","host"))
ggerrorplot(filter(ps_disease,trt=="C"|trt=="N1"|trt=="N5"|trt=="N10"|trt=="PK"|trt=="NPK"),x="trt",y="new_invert",desc_stat = "mean_sd",facet.by=c("date","host"))

#if plants are diseased, which ones get hit hardest?
ggerrorplot(filter(ps_disease,old_rust!=0 & (trt=="C"|trt=="N1"|trt=="N5"|trt=="N10"|trt=="PK"|trt=="NPK")),x="trt",y="old_rust",desc_stat = "mean_se",facet.by=c("date","host"))
```

## Specific questions
```{r}
#for treatments of interest (C, N1, N5, N10, PK, NPK), do we see any trends in rust scoring on old and new leaves for Andro and Agro at different points in the growing season? 
#graph scoring info across N added with separate lines for PK and no PK
#summarize by treatment, host species, and collection date

library(tidyverse)
library(magrittr)
library(ggplot2)

#add another column for infected individuals for old and new rust so we can get proportion infected for each treatment and sampling day
ps_disease_pk %<>% 
  mutate(old_rust_inf = case_when(
    old_rust>0 ~ '1',
    old_rust==0 ~ '0',
    TRUE ~ "NA"
  ),
  new_rust_inf = case_when(
    new_rust>0 ~ '1',
    new_rust==0 ~ '0',
    TRUE ~ "NA"),
  old_rust_inf=as.numeric(old_rust_inf),
  new_rust_inf=as.numeric(new_rust_inf))

#summary statistics
ps_disease_summary <- ps_disease_pk %>% 
  filter(trt=="C"|trt=="N1"|trt=="N5"|trt=="N10"|trt=="PK"|trt=="NPK") %>% 
  group_by(trt,date,host) %>% 
  summarize(mean_old_rust = mean(old_rust,na.rm=T),
            se_old_rust = sd(old_rust,na.rm=T)/sqrt(n()),
            mean_new_rust = mean(new_rust,na.rm=T),
            se_new_rust = sd(new_rust,na.rm=T)/sqrt(n()),
            N_added = unique(N_added),
            PK_added = unique(PK_added),
            )
#overall summary
#ps_disease_summary$trt %>% factor(ps_disease_summary$trt,levels=c("C","N1","N5","N10","PK","NPK"))
ps_disease_summary %>% ggplot(aes(x=date,y=mean_old_rust,color=trt,group=trt)) + geom_point() +geom_line()+facet_wrap(.~host+PK_added) + theme_bw(base_size=24)+geom_pointrange(aes(ymin=mean_old_rust-se_old_rust,ymax=mean_old_rust+se_old_rust))

#summary for proportion infected
ps_disease_summary_old_rust <- ps_disease_pk %>%
  filter(trt=="C"|trt=="N1"|trt=="N5"|trt=="N10"|trt=="PK"|trt=="NPK") %>%
  filter(!is.na(old_rust_inf)) %>% 
  group_by(trt,host,date) %>%
  summarize(old_rust_inf_count = sum(old_rust_inf),
            old_rust_inf_scored = n(),
            prop_old_rust_inf = old_rust_inf_count/old_rust_inf_scored,
            N_added = unique(N_added),
            PK_added = unique(PK_added))
  
ps_disease_summary_old_rust$trt <- factor(ps_disease_summary_old_rust$trt,levels=c("C","N1","N5","N10","PK","NPK"))
       
ps_disease_summary_old_rust %>% ggplot(aes(x=trt,y=prop_old_rust_inf)) + geom_point() + theme_bw() +facet_grid(cols=vars(date),rows=vars(host))

ps_disease_summary_old_rust %>% ggplot(aes(date,prop_old_rust_inf,color=host,group=host)) + geom_point() + theme_bw() + facet_wrap(.~trt,nrow=1) + geom_line()

ps_disease_summary_old_rust %>% filter(host=="Ange") %>% ggplot(aes(date,prop_old_rust_inf,color=trt,group=trt)) + geom_point() + theme_bw(base_size=24) + facet_wrap(.~PK_added) + geom_line()

#summary for proportion infected
ps_disease_summary_new_rust <- ps_disease_pk %>%
  filter(trt=="C"|trt=="N1"|trt=="N5"|trt=="N10"|trt=="PK"|trt=="NPK") %>%
  filter(!is.na(new_rust_inf)) %>% 
  group_by(trt,host,date) %>%
  summarize(new_rust_inf_count = sum(new_rust_inf),
            new_rust_inf_scored = n(),
            prop_new_rust_inf = new_rust_inf_count/new_rust_inf_scored,
            N_added = unique(N_added),
            PK_added = unique(PK_added))
  
ps_disease_summary_new_rust$trt <- factor(ps_disease_summary_new_rust$trt,levels=c("C","N1","N5","N10","PK","NPK"))
       
ps_disease_summary_new_rust %>% ggplot(aes(x=trt,y=prop_new_rust_inf)) + geom_point() + theme_bw() +facet_grid(cols=vars(date),rows=vars(host))

ps_disease_summary_new_rust %>% filter(host=="Ange") %>% ggplot(aes(date,prop_new_rust_inf,color=trt,group=trt)) + geom_point() + theme_bw(base_size=24) + facet_wrap(.~PK_added) + geom_line()
```



```{r}
#trying to get disease-y-ness per plot

cover %<>% filter(taxon=="ANDROPOGON GERARDII"|taxon=="AGROPYRON REPENS")
#julian days of sampling: 151, 155, 169, 189, 191, 204, 205

cover %<>%
  mutate(month= case_when(
    julian_day==151 ~ "June",
    julian_day==155 ~ "June",
    julian_day==169 ~ "bonus",
    julian_day==189 ~ "July",
    julian_day==191 ~ "July",
    julian_day==204 ~ "August",
    julian_day==205 ~ "August"
  )) %>%
  filter(month!="bonus")

#filter out bonus sampling later

#ps_sampling: 156, 189, 218

#summary of prop inf by plot
ps_disease_summary_for_cover <- ps_disease_pk %>%
  filter(trt=="C"|trt=="N1"|trt=="N5"|trt=="N10"|trt=="PK"|trt=="NPK") %>%
  filter(!is.na(old_rust_inf)) %>% 
  group_by(plot,host,month) %>%
  summarize(old_rust_inf_count = sum(old_rust_inf),
            old_rust_inf_scored = n(),
            prop_old_rust_inf = old_rust_inf_count/old_rust_inf_scored,
            N_added = unique(N_added),
            PK_added = unique(PK_added))
ps_disease_summary_for_cover2 <- ps_disease_pk %>%
  filter(trt=="C"|trt=="N1"|trt=="N5"|trt=="N10"|trt=="PK"|trt=="NPK") %>%
  filter(!is.na(new_rust_inf)) %>% 
  group_by(plot,host,month) %>%
  summarize(new_rust_inf_count = sum(new_rust_inf),
            new_rust_inf_scored = n(),
            prop_new_rust_inf = new_rust_inf_count/new_rust_inf_scored,
            N_added = unique(N_added),
            PK_added = unique(PK_added))
ps_disease_summary_for_cover %<>% mutate(count=old_rust_inf_count,
                                         scored=old_rust_inf_scored,
                                         prop_inf=prop_old_rust_inf,
                                         leaf="old") %>%
  select(-c("old_rust_inf_count","old_rust_inf_scored","prop_old_rust_inf"))

ps_disease_summary_for_cover2 %<>% mutate(count=new_rust_inf_count,
                                         scored=new_rust_inf_scored,
                                         prop_inf=prop_new_rust_inf,
                                         leaf="new") %>%
  select(-c("new_rust_inf_count","new_rust_inf_scored","prop_new_rust_inf"))

disease_cover<-bind_rows(ps_disease_summary_for_cover,ps_disease_summary_for_cover2)
cover%<>%select(-c("date","julian_day","block","subplot","functional.group","notes"))
cover %<>% mutate(host= case_when(
  taxon=="ANDROPOGON GERARDII" ~ "Ange",
  taxon=="AGROPYRON REPENS" ~ "Agre"
)) %>% select(-taxon)
  
cover %<>% unite(col="ID",c("host","month","plot"))
disease_cover %<>% unite(col="ID",c("host","month","plot"))
combo<-right_join(disease_cover,cover,by="ID")

combo %<>% filter(!is.na(PK_added)) %>% mutate(disease_burden = prop_inf*cover) %>% separate(col="ID",into=c("host","month","plot"),sep="_")
combo %>% ggplot(aes(x=N_added,y=disease_burden,color=trt,group=trt))+geom_boxplot()+facet_wrap(.~host+PK_added) + theme_bw()

combo %>% filter(month=="June") %>% ggplot(aes(x=N_added,y=disease_burden,color=PK_added,group=PK_added))+geom_point()+facet_wrap(.~host+leaf)+theme_bw()+geom_smooth(se=FALSE,method=lm)
combo %>% filter(month=="June") %>% ggplot(aes(x=N_added,y=cover,color=PK_added,group=PK_added))+geom_point()+facet_wrap(.~host+leaf)+theme_bw()+geom_smooth(se=FALSE,method=lm)



combo %>% filter(month=="July") %>% ggplot(aes(x=N_added,y=disease_burden,color=PK_added,group=PK_added))+geom_point()+facet_wrap(.~host)+theme_bw()+geom_smooth(se=FALSE,method=lm)
combo %>% filter(month=="August") %>% ggplot(aes(x=N_added,y=disease_burden,color=PK_added,group=PK_added))+geom_point()+facet_wrap(.~host)+theme_bw()+geom_smooth(se=FALSE,method=lm)

combo %>% ggplot(aes(x=N_added,y=cover,color=PK_added,group=PK_added))+geom_point()+facet_wrap(.~host)+theme_bw()+geom_smooth(se=FALSE,method=lm)
```


















