---
title: "nutnet disease"
author: "Katie Schroeder"
date: "2023-01-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## loading in
```{r loading}
#load in data
ks_disease<-read.csv("Stomata_Data_REU_2019_master.csv")
ps_disease<-read.csv("nutnet_fungus_2019_master.csv")
cover<-read.csv("NutNet_Species_Cover_May30_thru_August.csv")

library(tidyverse)
```

## Basics of each data frame
ks_disease = disease scoring by Megan and Katie for 2019 REU project.
plots: 13 - 24
treatments: all trts of block 2
dates sampled: "6.18.19"   "6.21.2019" "7.10.2019" "7.31.2019"
plants scored:"AGROPYRON REPENS"    "ANDROPOGON GERARDII" "POA PRATENSIS"

ps_disease = disease scoring by Rachel Paseka and Alex Strauss in summer 2019. Lots of manual labor by Evelyn and Hajira and maybe Sam?
plots: 1-60
treatments: all trts all blocks
dates sampled: "6/5/2019" "7/8/2019" "8/6/2019"
plants scored: "Agropyron repens"    "Andropogon gerardii"

cover = cover data by Megan and Katie for 2019 field season.
plots: 1-36, 39, 43, 44, 45, 48, 50, 55, 56, 58, 59
dates: "5.30.2019" "6.18.2019" "6.4.2019"  "7.10.2019" "7.23.2019" "7.24.2019" "7.8.2019" 
treatments: All for blocks 1-3, C, N1, N5, N10 for blocks 4 and 5

## Summaries
```{r summarize}
ps_disease %>%
  count(date,host,plot)
#1-20 replicates per plot per sampling day (most 1-20). Plants were flagged and tracked through time
ps_disease %>%
  count(date,host,trt)
#40-80 replicates per trt per sampling day

ks_disease %>% 
  filter(!is.na(fungal)) %>%
  count(date,plant,trt)
#5-7 replicates per plot per sampling day
```

## General trends
```{r}
#make a total fungal column in alex and rachel's data
ps_disease$fungal_old <- ps_disease$old_rust + ps_disease$old_bs
ps_disease$fungal_new <- ps_disease$new_rust + ps_disease$new_bs

library(ggpubr)
hist(ps_disease$old_rust,breaks=50)
hist(ps_disease$old_bs,breaks=50)

hist(ps_disease$new_rust,breaks=50)
#old rust is most frequently not 0

ps_disease$trt<-factor(ps_disease$trt,levels=c("C","C_H","N1","N5","N10","P","K","PK","NK","NP","NPK","NPK_H"))
ggerrorplot(ps_disease,x="trt",y="old_rust",desc_stat = "mean_sd",facet.by=c("date","host"))
```

